### 2025-11-08 thesis/implementation/index.tex
Use appendix

As outlined above the \textsc{OmniSQL} prompt was used as base for
\textsc{Natural}'s prompt together with a code representation prompt.
\textsc{Natural} uses more explicit instructions compared to DAIL-SQL. Key
differentiations include the inclusion of similarity scores to give the model
the ability to weight the samples itself.

@@

\begin{verbatim}
EXAMPLE OF THE PROMPT
\end{verbatim}

@@

The refinement prompt uses a structured format similar to ICL generation but
with key differences: it includes the original question as context, provides
the candidate query that needs fixing, and uses explicit correction language.

\begin{verbatim}
    SHOW PROMPT LAYOUT
\end{verbatim}

@@

The vector database implementation \texttt{Vector} is used both during
sampling to construct the vector db and runtime to run the sample selection
algorithm. Using SQLite and \texttt{sqlite-vec} to implement cosine search in
text corpora of up to a few million samples dramatically simplified the
development of \textsc{Natural} as no manual performance optimizations had to
be implemnented. SQLite has a second architectural benefit besides relatively
good performance, which is the portability characteristics of having all
samples indexed in a single file on disk. This enables swapping out the
respectively used samples by using another \texttt{.vector} file on disk.

\begin{verbatim}
let question = "How many ..";

// Initialize vector database for similarity search
let vector = Vector::infer();
let model = EmbeddingModel::from_env().expect("failed to load embedding model");

// Query distance indices
let index = DistanceIndexRow::get("test_db", &vector)?;

// Embed question and perform similarity search
let selector = Selector::new(&question, None, &model, &vector)?;
let samples = selection::<4>(
    &Selector::new(
        question.as_ref(),
        None,
        &model,
        &vector,
    )?,
    &index.into(),
    &vector,
)?;

\end{verbatim}

@@

As the \texttt{.vector} file is still a regular SQLite database, alongside
indexing tables using embeddings (\texttt{float[4096]} or \texttt{float[2048]})
regular database tables can be maintained to save data derived at sampling
time. \texttt{Vector} hosts the following database schema:

\begin{verbatim}
CREATE TABLE IF NOT EXISTS whitelist (
  word             text not null primary key,
  frequency        integer not null,
  domain_count     integer not null
);

CREATE TABLE IF NOT EXISTS databases (
  id               text not null,
  name             text not null,
  graph            text not null
);

CREATE TABLE IF NOT EXISTS distance_indices (
  db_id            text not null,
  distances        text not null
);

CREATE VIRTUAL TABLE IF NOT EXISTS samples USING vec0 (
  id                 text not null,
  db_id              text not null,

  question           text not null,
  question_masked    text not null,
  question_embedding float[2048],

  sql                text not null,
  sql_masked         text not null,
  sql_embedding      float[2048]
);
\end{verbatim}

>>>>

As outlined above the \textsc{OmniSQL} prompt was used as base for
\textsc{Natural}'s prompt together with a code representation prompt.
\textsc{Natural} uses more explicit instructions
(see~\ref{appendix:inference-prompt}) compared to DAIL-SQL. Key
differentiations include the inclusion of similarity scores to give the model
the ability to weight the samples itself.

@@

@@

The refinement prompt uses a structured format similar to ICL generation but
with key differences: it includes the original question as context, provides
the candidate query that needs fixing, and uses explicit correction language
(see \ref{appendix:refinement-prompt}).

@@

The vector database implementation \texttt{Vector} is used both during
sampling to construct the vector db and runtime to run the sample selection
algorithm. Using SQLite and \texttt{sqlite-vec} to implement cosine search in
text corpora of up to a few million samples dramatically simplified the
development of \textsc{Natural} as no manual performance optimizations had to
be implemnented. SQLite has a second architectural benefit besides relatively
good performance, which is the portability characteristics of having all
samples indexed in a single file on disk. This enables swapping out the
respectively used samples by using another \texttt{.vector} file on disk.
See \ref{appendix:vector-db-api} for reference.

@@

As the \texttt{.vector} file is still a regular SQLite database, alongside
indexing tables using embeddings (\texttt{float[4096]} or \texttt{float[2048]})
regular database tables can be maintained to save data derived at sampling
time. \texttt{Vector} hosts the database schema described
in \ref{appendix:vector-schema}.
